---
title: "MODIS Tiles in Leaflet"
author: "Anna Vasylytsya"
date: "01/09/2020"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
proj_path <- "C:/Users/Anna V/Documents/GitHub/modis_tiles"
```

```{r libraries}
library(MODIS)
library(rgdal)
library(leaflet)
library(sp)
library(printr)
```

## Get MODIS Tiles from MODIS package
MODIS provided tile calculator: https://landweb.modaps.eosdis.nasa.gov/cgi-bin/developer/tilemap.cgi
```{r}
sa <- readOGR("C:/Users/Anna V/Documents/GitHub/exploring_spatial_data/data/south_america_shp/SouthAmerica.shp")
gtm <- sa[sa@data$Name== "GUATEMALA", ]
```

```{r MODIS}
modis_tile <- getTile(extent(gtm))
modis_tile@tile
```

## Load shapefile

```{r load_shp}
modis <- readOGR(paste0(proj_path, "/data/modis_sinusoidal_grid_world.shp"), 
                 verbose = FALSE, stringsAsFactors = FALSE)
print(str(modis@data))
print(head(modis@data, n=3)) # the combination of h and v, indicates the modis tile
plot(modis, main = "MODIS tile grid")
```

The tiles are square! The whole grid is a rectangle! "There are 460 non-fill tiles, tiles are 10 degrees by 10 degrees at the equator." https://modis-land.gsfc.nasa.gov/MODLAND_grid.html

### Coordinate reference system (CRS)
```{r crs}
print((modis@proj4string)) # modis sinusoidal projection!
```

Change the string to numbers
```{r numeric}
modis$h <- as.numeric(modis$h)
modis$v <- as.numeric(modis$v)
```

## Generate a tile label within the modis dataframe 
Modis tiles are specified as: h##v##, this means that we will have to pad the single digit numbers with zeros.
```{r padding_function}
lead0 <- function(m, pad.val){
    # A function to pad values with a specified characters
    # Only designed to handle up to two characters
    # Args:
    #       m: a vector
    #       pad.val: the padding character (e.g. "0")
    # Returns:
    #       A vector where all elements have two characters
  m[nchar(m)==1] <- paste0(pad.val,m[nchar(m)==1])
  return(m)
}
```

```{r label}
modis$label <- paste0("<b>h</b>", lead0(modis$h, 0), "<b>v</b>", lead0(modis$v, 0))
print(head(modis@data, n=3))
```

## Project modis tiles shapefile to WGS84
"The Leaflet package expects all point, line, and shape data to be specified in latitude and longitude using WGS 84 (a.k.a. EPSG:4326). By default, when displaying this data it projects everything to EPSG:3857 and expects that any map tiles are also displayed in EPSG:3857."
"While tiles must be in the same projection as used in the leafletCRS function, you must always use WGS 84 longitude/latitude data for markers, circles, polygons, and lines. Leaflet will automatically project the coordinates when displaying."

```{r pop_rast, include = FALSE}
# pop <- raster(paste0(gsub("modis_tiles", "", proj_path), "/exploring_spatial_data/data/pop17_tif/ppp_2017_1km_Aggregated.tif"))
```

```{r modis_wgs84}
# maybe try rasterizing the polygon, changing the CRS, then rasterToPolygon and plotting! av.note: 2020-01-21
#https://stackoverflow.com/questions/26102012/plotting-world-map-in-orthographic-projection-is-giving-non-finite-points/26153202#26153202

modis_wgs84 <- spTransform(modis, CRS("+init=EPSG:4326"))
print(modis_wgs84@proj4string)
str(modis_wgs84@data)

m84_rast <- rasterize(modis_wgs84, pop, updateValue = TRUE)
```

```{r epsg:4857, include = FALSE}
# bol_pop <- raster(paste0(gsub("modis_tiles", "", proj_path), "/exploring_spatial_data/data/bolivia/bolivia_ppp_2017_1km.tif")
# bol_web <- projectRasterForLeaflet(x = bol_pop, method = "ngb")
leaflet_crs <- "+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +no_defs"
# modis_lf <- spTransform(modis_wgs84, CRS(leaflet_crs)) #not possible to project the whole world
```

```{r write_modis_wgs84}
writeOGR(modis_wgs84, paste0(proj_path, "/data/modis_grid_wgs84.shp"), 
         layer = "/modis_grid_wgs84.shp", driver = "ESRI Shapefile")
```

## Interactive Map

```{r leaflet}
#inspiration: https://rpubs.com/bhaskarvk/proj4leaflet
leaflet(data = modis, options = 
          leafletOptions(crs=leafletCRS(crsClass="L.Proj.CRS",
                                        code='SR-ORG:6842',
                                        proj4def= "+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs",
                                        resolutions = c(65536, 32768, 16384, 8192, 4096, 2048)))) %>% # i think i have to play with the resolutions!
  # setView(lng = -71.0589, lat = 42.3601, zoom = 2) %>% #
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(color = "black", opacity = 1, weight = 1, popup = modis$label,
              fill = TRUE,fillColor = "white", fillOpacity = 0.2
              #smoothFactor = 0.5, group = "tile"
              )
```

If that map look quite weird to you, please read this blog post that breaks down coordinate reference systems and how 3D data is displayed in two dimensions: https://lyzidiamond.com/posts/4326-vs-3857